---
title: 'Final Project 2: Reproducible Report on COVID19 Data"'
author: "TSP"
date: "11/30/2022"
output:
  pdf_document: default
  word_document: default
---

#Peer-graded Assignement: NYPD Shooting Incident Data Report

Assignement Tasks:
Import, tidy and analyze the COVID19 dataset from the Johns Hopkins github site. This is the same dataset I used in class. Feel free to repeat and reuse what I did if you want to. Be sure your project is reproducible and contains some visualization and analysis that is unique to your project. You may use the data to do any analysis that is of interest to you. You should include at least two visualizations and one model.  Be sure to identify any bias possible in the data and in your analysis.


#Step 1: Install packages and enable the package required for data analysis
```{r include = FALSE}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(usmap)
library(zoo)
```

#Step 2: Import data from COVID-19 Data Repository by the Center for Systems Science and Engineering (CSSE) 
# at Johns Hopkins University
```{r setup, include=TRUE, cache = FALSE}
url_in <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/"
file_names <- c("time_series_covid19_confirmed_global.csv",
                 "time_series_covid19_deaths_global.csv",
                 "time_series_covid19_confirmed_US.csv",
                 "time_series_covid19_deaths_US.csv")
urls <- str_c(url_in,file_names)
global_cases <- read_csv(urls[1])
global_deaths <- read_csv(urls[2])
US_cases <- read_csv(urls[3])
US_deaths <- read_csv(urls[4])
```


#Step 3: Clean and Tidy  data
```{r tidy_data}
#Tidy  global data
global_cases <- global_cases %>%
  pivot_longer(cols = -c(`Province/State`,
                         `Country/Region`,Lat,Long),
               names_to = "date",
               values_to = "cases") %>%
  select(-c(Lat,Long))

global_deaths <- global_deaths %>%
  pivot_longer(cols = -c(`Province/State`,
                         `Country/Region`,Lat,Long),
               names_to = "date",
               values_to = "deaths") %>%
  select(-c(Lat,Long))

global <- global_cases %>%
  full_join(global_deaths) %>%
  rename(Country_Region = `Country/Region`,
         Province_State = `Province/State`) %>%
  mutate(date = mdy(date))

#Check ddescriptive statitic and information 
summary(global)

# Remove the zero case 
global <- global %>% filter(cases > 0)
summary(global)

# Check maximum case whether it is correct or not
global <- global %>% filter(cases > 28000000)
global

#Tidy US data
US_cases
US_cases <- US_cases %>%
  pivot_longer(cols = -(UID:Combined_Key),
               names_to = "date",
               values_to = "cases") %>%
  select(Admin2:cases) %>%
  mutate(date = mdy(date)) %>%
  select(-c(Lat,Long_))

US_cases

US_deaths
US_deaths <- US_deaths %>%
  pivot_longer(cols = -(UID:Population),
               names_to = "date",
               values_to = "deaths") %>%
  select(Admin2:deaths) %>%
  mutate(date = mdy(date)) %>%
  select(-c(Lat,Long_))
US_deaths

US <- US_cases %>%
  full_join(US_deaths)
US

#Add population to global data with look up table
global <- global %>%
  unite("Combined_Key",
        c(Province_State, Country_Region),
        sep = ", ",
        na.rm = TRUE,
        remove = FALSE)

#get uid lookup url
uid_lookup_url <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/UID_ISO_FIPS_LookUp_Table.csv"
uid <- read_csv(uid_lookup_url)%>%
  select(-c(Lat, Long_, Combined_Key, code3, iso2, iso3, Admin2))

#Join look up table with global
global <- global %>%
  left_join(uid, by = c("Province_State", "Country_Region")) %>%
  select(-c(UID, FIPS)) %>%
  select(Province_State, Country_Region, date, cases,deaths,Population,Combined_Key
         )
global

```
#Step 4: Visualize
```{r Visualization}
#Transform data 
US_by_state <- US %>%
  group_by(Province_State, Country_Region,date) %>%
  summarize(cases = sum(cases), deaths = sum(deaths),
            Population = sum(Population)) %>%
  mutate(deaths_per_mil = deaths * 1000000/Population) %>%
  select(Province_State,Country_Region,date,
         cases,deaths, deaths_per_mil,Population) %>%
  ungroup()

US_by_state

US_totals <- US_by_state %>%
  group_by(Country_Region, date) %>%
  summarize(cases = sum(cases), deaths = sum(deaths),
            Population = sum(Population)) %>%
  mutate(deaths_per_mil = deaths * 1000000 / Population) %>%
  select(Country_Region,date,
         cases,deaths,deaths_per_mil,Population) %>%
  ungroup()
US_totals
tail(US_totals)

#Perform data visualization - Visualization_1
# Total Covid case by time 
US_totals %>% 
  filter(cases > 0)%>%
  ggplot(aes(x = date, y=cases)) +
  geom_line(aes(color = "Cases")) +
  geom_point(aes(color = "cases")) +
  geom_line(aes(y = deaths, color = "deaths")) +
  geom_point(aes(y = deaths, color = "deaths")) +
  scale_y_log10() +
  theme(legend.position = " bottom",
        axis.text.x = element_text(angle = 90))+
  labs(title = "COVID-19 in US", y = NULL)

#Perform data visualization - Visualization_2
# New Yosk Covid cases by time
state <- "New York"
US_by_state %>% 
  filter(Province_State == state) %>%
  filter(cases > 0) %>%
  ggplot(aes(x = date, y=cases)) +
  geom_line(aes(color = "cases")) +
  geom_point(aes(color = "cases"))+
  geom_line(aes(y = deaths, color = "deaths"))+
  geom_point(aes(y = deaths, color = "deaths"))+
  scale_y_log10() +
  theme(legend.position = " bottom",
        axis.text.x = element_text(angle = 90))+
  labs(title = str_c("COVID-19 in ",state), y = NULL)
```

#Step 4: Analyzing
```{r Analysis }
#Transform data - add new_cases and new deaths columns
US_by_state <- US_by_state %>%
  mutate(new_cases = cases - lag(cases),
         new_deaths = deaths - lag(deaths))
US_totals <- US_totals %>%
  mutate(new_cases = cases - lag(cases),
         new_deaths = deaths - lag(deaths))
#checking data
tail(US_by_state)
tail(US_totals)
tail(US_totals %>% select(new_cases,new_deaths,everything()))

#Graph US_total with new_cases and new deaths
US_totals %>% 
  ggplot(aes(x = date, y = new_cases)) +
  geom_line(aes(color = "new_cases"))+
  geom_point(aes(color = "new_cases")) +
  geom_line(aes(y = new_deaths, color = "new_deaths")) +
  geom_point(aes(y = new_deaths, color = "new_deaths")) +
  scale_y_log10()+
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90)) +
  labs(title = "COVID19 in US", y=NULL)

# Find top-ten state smallest deaths in thousand
US_state_totals <- US_by_state %>%
  group_by(Province_State) %>%
  summarize(deaths = max(deaths), cases = max(cases),
            population = max(Population),
            cases_per_thou = 1000 * cases/ population,
            deaths_per_thou = 1000 * deaths / population) %>%
  filter(cases > 0, population > 0)
US_state_totals %>%
  slice_min(deaths_per_thou, n= 10)%>%
  select(deaths_per_thou, cases_per_thou, everything())

# Find top-ten largest deaths in thousand 
US_state_totals %>%
  slice_max(deaths_per_thou, n=10) %>%
  select(deaths_per_thou, cases_per_thou, everything())
```


#Step 4: Modelling
```{r Modelling}
#Linear Regression Model 
mod <- lm(deaths_per_thou ~ cases_per_thou, data = US_state_totals)
summary(mod)

x_grid <- seq(1,151)
new_df <- tibble(cases_per_thou = x_grid)
US_state_totals %>% mutate(pred = predict(mod))

US_tot_w_pred <- US_state_totals %>% mutate(pred = predict(mod))
US_tot_w_pred

# Visualize it
US_tot_w_pred %>% ggplot() +
  geom_point(aes(x = cases_per_thou, y= deaths_per_thou), color = "blue") +
  geom_point(aes(x = cases_per_thou, y = pred), color = "red")
```


#Step 5: My two unique visualizations and one model 
```{r My two unique visualizations and one model}

#Produce map to present  number of case and deaths.
mapdata <- map_data("world")
view(mapdata)
mapdata <- left_join(mapdata, global, by=c('region'='Country_Region'))
mapcases <- ggplot(mapdata, aes(x = long, y= lat, group=group)) +
  geom_polygon(aes(fill = cases), color = 'black')
mapdeaths <- ggplot(mapdata, aes(x = long, y= lat, group=group)) +
  geom_polygon(aes(fill = deaths), color = 'black')

#Produce map to present deaths_per_mil in US. 
colnames(US_by_state)[1] <- "state"
plot_usmap(data = US_by_state, values = "deaths_per_mil", color = "red") + 
  scale_fill_continuous(name = "Population (2015)", label = scales::comma) + 
  theme(legend.position = "right")

#Develop the model
global_group <- global %>% 
  group_by(date) %>%
  mutate(deaths_per_thou = deaths / 1000) %>%
  select(date,deaths_per_thou)
#develop simple time-series model
tsmodel <- ts(global_group$deaths_per_thou,start = c(2018,2,8),frequency = 365)
#Plot the result
plot(tsmodel)

```